---
title: "Assignment 7 - Working with XML, JSON, and HTML in R"
author: "Stefano Biguzzi"
date: "10/9/2020"
output:
  html_document:
    includes:
      in_header: header.html
    css: ./project.css
    highlight: pygments
    theme: cerulean
    toc: false
    code_folding: show
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, comment=NA, warning=F)
```

# Loading Libraries
```{r,message=F}
library(tidyverse)
library(RCurl)
library(XML)
library(rjson)
library(knitr)
```

# Introduction
We were asked to create a XML file, a Json file, and a Html file. We were then asked to load them into R and create a data frame from the data. I decided to use the `tidyverse` and `Rcurl` libraries for loading in data,`XML` for loading in the XML and Html files,`rjson` to load in the json file, and `knitr` for table purposes, Below are three tabs each explaining how to load in one of the files and transform it into a data frame.

# How to Extract Data from different file types{.tabset .tabset-fade}

## 1. Extracting data from XML file  

1. Use **`xmlParse`** to read in the file as an XML class document
```{r}
rawXML <-
  xmlParse(getURL("https://raw.githubusercontent.com/sbiguzzi/data607assignment7/main/book-list-xml.xml"))
```
2. Then we use **`xmlRoot`** to allow R to interact with the XML file so we can call the first node (element) of the file
```{r}
rootXML <- xmlRoot(rawXML)
```
When we call the first element of *rootXML* we see that we have the first book
```{r, echo=F}
rootXML[1]
```
3. Then we can use the **`xmlSApply`** function to iterate through each *header* and *value* within the XML node and return a matrix.
```{r}
dataXML <- xmlSApply(rootXML, function(x) xmlSApply(x, xmlValue))
```
```{r,echo=F}
kable(dataXML)
```
***Note:*** \*If we only used `xmlSApply` once, R only reads the book header and then the values, as seen below, rather than reading both the headers and sub-headers
```{r, echo=F, out.width="75%"}
xmlSApply(rootXML,xmlValue)
```

4. Finally we use the **`data.frame`** function to transpose the matrix and convert it to a data frame.
```{r}
book.xml.df <- data.frame(t(dataXML),row.names=NULL)
```
```{r, echo=F}
kable(book.xml.df)
```

## 2. Extracting data from JSON file  

1. First load the Json file using the **`fromJSON`** function.
```{r}
rawJSON <- 
  fromJSON(getURL("https://raw.githubusercontent.com/sbiguzzi/data607assignment7/main/book-list-json.json"))
```

2. Create a table that stores the book list data as the row
```{r}
tibJSON <- rawJSON %>%
  map_if(is.data.frame,list) %>%
  as_tibble() %>%
  unnest(cols = c('book1', 'book2', 'book3'))
```

3. Rename the rows and transpose the data frame
```{r}
#Transpose data
book.json.df <- as.data.frame(t(as.matrix(tibJSON)))

#Rename columns
names(book.json.df) <-
  c("title","author1","author2","genre","year")

#Reset row names
rownames(book.json.df) <- NULL
```
```{r, echo=F}
kable(book.json.df)
```

## 3. Extracting data from HTML file 

1. Load the data
```{r}
rawHTML <- getURL("https://raw.githubusercontent.com/sbiguzzi/data607assignment7/main/book-list-html.html")
```

2. Write it to a data frame and rename columns
```{r}
#Create data frame
book.html.df <- as.data.frame(readHTMLTable(rawHTML, stringAsFactors = FALSE))

#Rename columns
names(book.html.df) <-
  tolower(gsub("NULL.","",names(book.html.df)))
```
```{r, echo=F}
kable(book.html.df)
```

# Conclusion
All three of the files can be read into R and transformed into the same data frame, see table below, using different functions and libraries. However, it would seem that HTML is the easiest file format to read in as it does not require any transformation of the data. This is because with both Json and XML files you have to read in the data and tell R what the column names and values are, whereas with HTML there is code to tell R which is which.

```{r, echo=F}
kable(list(
  as.data.frame(sapply(book.xml.df, class)),
  as.data.frame(sapply(book.json.df, class)),
  as.data.frame(sapply(book.html.df, class))), caption = "Compare data types for columns in each dataframe")
```



***

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
